
**Machine Learning lab 4**

Kucheriavyi Yaroslav. ІАВ-21

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(lattice); library(ggplot2); library(caret); library(randomForest); library(rpart); library(rpart.plot);
```

```{r}
# data upload and cleanup
trainingset<-read.csv("C:/Users/NinjaWizard/Documents/pml-training.csv",na.strings=c("NA","#DIV/0!", ""))
testingset<-read.csv("C:/Users/NinjaWizard/Documents/pml-testing.csv",na.strings=c("NA","#DIV/0!", ""))
original_testingset<-testingset
original_testingset$X<-NULL

# Clear missing values
trainingset<-trainingset[,colSums(is.na(trainingset)) == 0]
testingset<-testingset[,colSums(is.na(testingset)) == 0]

# Clear irrelevant columns
trainingset<-trainingset[,-c(1:7)]
testingset<-testingset[,-c(1:7)]

# Split training/testing data
traintrainset<-createDataPartition(y=trainingset$classe,p=0.75,list=FALSE)
TrainTrainingSet<-trainingset[traintrainset, ] 
TestTrainingSet<-trainingset[-traintrainset, ]
```

**Prediction model 1: Decision Tree **

```{r}
model1<-rpart(
    classe ~ .,
    data=TrainTrainingSet,
    method="class"
)

prediction1<-predict(model1,TestTrainingSet,type="class")

# Plot the Decision Tree
rpart.plot(model1,main="Classification Tree",extra=102,under=TRUE,faclen=0)

# Test results on our TestTrainingSet data set:
confusionMatrix(prediction1,as.factor(TestTrainingSet$classe))
```

**Prediction model 2: Random Forest**

```{r}
TrainTrainingSet$classe = factor(TrainTrainingSet$classe)
model2<-randomForest(
    classe ~. ,
    data=TrainTrainingSet,
    method="class"
)

# Predicting:
prediction2<-predict(model2,TestTrainingSet,type="class")

# Test results on TestTrainingSet data set:
confusionMatrix(prediction2,as.factor(TestTrainingSet$classe))
```

**Result: ** Random forest algorithm shows better performance of 99.57% accuracy versus 74.31% of desicion tree. Thus, random forest is considered superior for the task given. 

The final outcome of random forest applied against the testing dataset is stored as `result.csv`

```{r}
# predict outcome levels on the original Testing data set using Random Forest algorithm
predictfinal<-predict(model2,testingset,type="class")
original_testingset$classe<-predictfinal
result<-original_testingset[,c(ncol(original_testingset),1:(ncol(original_testingset)-1))]
write.csv(result,'./result.csv')
```